material and methods

### 2018-02-19 analysis/drought/explore_prec_longterm.Rmd
update

```{r packages, warning=FALSE, message=FALSE, echo=FALSE}
library("here")
library("tidyverse")
library("lubridate")
library("ggplot2")
library("ggrepel")
```

@@

>>>>

```{r packages, warning=FALSE, message=FALSE, echo=FALSE}
library("here")
library("tidyverse")
library("lubridate")
library("ggplot2")
library("ggrepel")
library("sp")
library("sf")
library("pander")
```

@@

# Mapa de estaciones 
```{r}
es <- read.csv(here::here('data_raw/meteo/prec_subclima/mdestaciones', 'estaciones_latlong.csv'), header=TRUE, sep=";")

es <- st_as_sf(x = es, 
               coords = c("Longitude","Latitude"), 
               crs = "+proj=longlat +datum=WGS84")

# mapview(es)

### 2018-02-19 man/ms_1_material_methods.Rmd
update

The populations of Pyrenean oak forests at Sierra Nevada are considered relict forests [@Melendo2000; @Vivero2000], undergoing intensive anthropic use in history [@CamachoOlmedo2002; @Valbuena2010; @Valbuena2017]. In fact, the status of conservation of this species for southern Spain is "Vulnerable" [@Vivero2000]. The relict presence of this species in Sierra Nevada is related both to its genetic resilience and high intraspecific genetic diversity [@Valbuena2010]. However, they are also expected to suffer the impact of climate change, due to their greater hydric requirements compared to other more drought-tolerant Mediterranean evergreen oak and pine species. Thus, simulations of the climate change effects on this habitat forecast a reduction in suitable habitats for Sierra Nevada [@Benito2011].

@@

The years 2005 and 2012 were chosen as drought years for this study because they have been documented as two of the worst drought years in the last decades for the southern Iberian Peninsula [@GarciaHerrera2007; @Gouveia2015; @Trigo2013] and they were charecterized as extreme drought in our climatic data (Figure S1). Aunque sabemos que existen otros años de sequía extrema (e.g. 1995, 1981) [@VicenteSerrano2014], no se han elegido por una limitación metodológica, esto es, nuestro objetivo es analizar el efecto de la sequía y la resiliencia a los eventos de sequía utilizando información de satélite y de dendrocronología, y la información de satélite utilizada (ver mas adelante) solamente está disponible desde el año 2000 en adelante.

@@

MODIS EVI Data from Collection 6 were obtanied using Google Earth Engine platform for the period 2000 - 2016. EVI data consits of 16-day maximun value composite images (23 per year) of the EVI value with a spatial resolution of 250 m x 250 m. We selected the pixels covering the distribution of *Q. pyrenaica* forests in Sierra Nevada (*n* = 928 pixels). A data filtering was applied to select EVI valid values. The filtering was done using quality flags and VI Usefulness Indices accompanying the EVI data. We filter out those values affected by high content of aerosols, clouds, snow and shadows, following the filtering recomendations of EVI data for mountain regions [@Reyes2015].

>>>>

The populations of Pyrenean oak forests at Sierra Nevada are considered relict forests [@Melendo2000; @Vivero2000], undergoing intensive anthropic use in history [@CamachoOlmedo2002; @Valbuena2010; @Valbuena2017]. In fact, conservation status of this species for southern Spain is "Vulnerable" [@Vivero2000]. The relict presence of this species in Sierra Nevada is related both to its genetic resilience and high intraspecific genetic diversity [@Valbuena2010]. However, they are also expected to suffer the impact of climate change, due to their greater hydric requirements compared to other more drought-tolerant Mediterranean evergreen oak and pine species. Thus, simulations of the climate change effects on this habitat forecast a reduction in suitable habitats for Sierra Nevada [@Benito2011].

@@

The years 2005 and 2012 were chosen as drought years for this study because they have been documented as two of the worst drought years in the last decades for the southern Iberian Peninsula [@GarciaHerrera2007; @Gouveia2015; @Trigo2013] and they were charecterized as extreme drought in our climatic data (Figure S1). Aunque sabemos que existen otros años de sequía extrema (e.g. 1995, 1981) [@VicenteSerrano2014], no se han elegido por una limitación metodológica, esto es, nuestro objetivo es analizar el efecto de la sequía y la resiliencia a los eventos de sequía utilizando información de satélite y de dendrocronología, y la información de satélite utilizada (ver mas adelante) solamente está disponible desde el año 2000 en adelante. We compare only these previous drought years in radial growth-time series, to place older drought events with those in 2005 and 2012 ... DUDA GUILLERMO

* ¿Decimos aquí algo así como?: Para los datos de radial-growth hemos computado la resiliencia en los 10 eventos mas severos (ver tabla S1)

@@

MODIS EVI Data from Collection 6 were obtanied using Google Earth Engine platform for the period 2000 - 2016. EVI data consits of 16-day maximun value composite images (23 per year) of the EVI value with a spatial resolution of 250 m x 250 m. We selected the pixels covering the distribution of *Q. pyrenaica* forests in Sierra Nevada (*n* = 928 pixels). A data filtering was applied to select EVI valid values. The filtering was done using quality flags and VI Usefulness Indices accompanying the EVI data. We filter out those values affected by high content of aerosols, clouds, snow and shadows, following the filtering recommendations of EVI data for mountain regions [@Reyes2015].
